---
title: "Data Analyst - Technical Assessment"
output: html_document
date: "2025-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (Optional) Install Setup Packages

```{r}
install.packages("readxl")
install.packages("janitor")
install.packages("dplyr")
install.packages("skimr")
```

## Load Test Dataset Into Memory

```{r}
library(readxl)
library(janitor)
library(dplyr)
library(skimr)


test_file <- "../../test-data.xlsx"

loans_raw <- read_xlsx(test_file, sheet = 1L) |>
    clean_names()
cashflows_raw <- read_xlsx(test_file, sheet = 2L) |>
    clean_names()
```

## Inspect Data

```{r}
skim(loans_raw)
```

```{r}
skim(cashflows_raw)
```

## Task 1 - Restructure Loan Data

The client has provided us with the loan data in a format that is not compatible
with our internal loan data schema. In particular, the data provided by the
client is a single table that contains the characteristics for both the loans
and the borrowers.
Our internal schema specifies that the loan and borrower details should belong
in separate tables with an additional relationship table linking them both based
on their respective primary ids.

The desired schema for the loan information is described below, where each
key-value pair represents the field name-type belonging to each table

```{r}
target_loan_schema <- list(
    loans = list(
        "contract_number" = "character", # PRIMARY KEY
        "termination_date" = "date",
        "termination_reason" = "character",
        "interest_rate" = "double",
        "loan_amount" = "double",
        "sum_of_last_payment" = "double",
        "debt_principal" = "double",
        "debt_interest" = "double",
        "total_outstanding_and_debt_amount" = "double",
        "proceeding_start_date" = "date",
        "delay_days_principal" = "integer"
    ),
    borrowers = list(
        "customer_id" = "character", # PRIMARY KEY
        "year_of_birth" = "integer",
        "gender" = "character",
        "number_of_customer_active_contracts" = "integer",
        "customer_active_contract_numbers" = "character"
    ),
    id_loan_borrower = list(
        "contract_number" = "character", # PRIMARY KEY
        "customer_id" = "character"
    )
)
```

## Task 1 - Solution

```{r}
library(purrr)

double_fields <- flatten(target_loan_schema) |>
    keep(\(val) val == "double") |>
    names()

char_fields <- flatten(target_loan_schema) |>
    keep(\(val) val == "character") |>
    names()

date_fields <- flatten(target_loan_schema) |>
    keep(\(val) val == "date") |>
    names()

integer_fields <- flatten(target_loan_schema) |>
    keep(\(val) val == "integer") |>
    names()

convert_double <- function(vec) {
    if (is.numeric(vec)) {
        return(as.double(vec))
    }
    as.double(gsub(",", ".", vec, fixed = TRUE))
}

loans <- loans_raw |>
    select(all_of(names(target_loan_schema$loans))) |>
    filter(!duplicated(.data$contract_number)) |> # Removes duplicate Ids
    mutate(
        across(any_of(char_fields), as.character),
        across(any_of(integer_fields), as.integer),
        across(any_of(date_fields), convert_to_date), # handles date type conversion
        across(any_of(double_fields), convert_double) # handles "," decimal separator
    )


borrowers <- loans_raw |>
    select(all_of(names(target_loan_schema$borrowers))) |>
    filter(!duplicated(.data$customer_id)) |> # Removes duplicate Ids
    mutate(
        across(any_of(char_fields), as.character),
        across(any_of(integer_fields), as.integer),
    )

id_loan_borrower <- loans_raw |>
    select(all_of(names(target_loan_schema$id_loan_borrower))) |>
    filter(!duplicated(.data$contract_number)) |> # Removes duplicate Ids
    mutate(
        across(any_of(char_fields), as.character),
    )
```


## Task 2 - Restructure Cashflow Date

Similar to the loan data, the client has provided the cash flow information
in a format that is incompatible with out internal cash flow schema.
Specifically, the client has provided the data in a "wide" format where each
cash flow date is a column. We require the data to be in a "long" format
according to the following schema

```{r}
target_cashflow_schema <- list(
    cashflows = list(
        "contract_number" = "character", # PRIMARY KEY - Component 1
        "cashflow_date" = "date", # PRIMARY KEY - Component 2
        "cashflow_amount" = "double"
    )
)
```

## Task 2 - Solution
```{r}
library(tidyr)
library(lubridate)

format_dates <- function(vec) {
    dates <- gsub("x", "", vec) |>
        paste0("_01") |>
        as.Date(format = "%Y_%m_%d") |>
        ceiling_date(unit = "months")
    dates - days(1)
}

cashflows <- cashflows_raw |>
    rename("contract_number" = "contract_nr") |> # Make loan id field name consistent
    pivot_longer(
        starts_with("x20"),
        names_to = "cashflow_date",
        values_to = "cashflow_amount"
    ) |>
    mutate(
        cashflow_date = format_dates(.data$cashflow_date) # converts to last day of corresponding months
    )
```


## Task 3 - Calulated Fields

Although the client data set contains all of the data that we require, it is
lacking certain fields that are needed by subsequent data processing pipelines.
In particular, we need to generate the following fields based on the existing
fields we have available:

1. Cumulative Cash Flow Amounts - For each loan we need to have a field that
   contains the sum of all cash flow amounts received to date.

2. Days Past Due (Buckets) - For each loan we require a field that classifies
   the loan based on whether the field `delay_days_principal` falls into one of
   the following buckets:

            [0, 180), [180, 360), [360, 540), [540, 720), 720+

## Task 3 - Solution
```{r}
cumulative_cashflow <- summarise(
    cashflows,
    cumulative_cashflow_amount = sum(cashflow_amount, na.rm = TRUE),
    .by = "contract_number"
)

loans |>
    left_join(cumulative_cashflow, by = "contract_number") |>
    mutate(
        days_past_due_bucket = cut(
            delay_days_principal,
            breaks = c(-Inf, 180, 360, 540, 720, Inf),
            labels = c(
                "[0, 180)", "[180, 360)", "[360, 540)", "[540, 720)", "720+"
            ),
            right = FALSE # Ensures the buckets are "closed" on the left
        )
    )
```
