---
title: "Data Analyst - Technical Assessment"
output: html_document
date: "2025-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (Optional) Install Setup Packages

```{r}
install.packages("readxl")
install.packages("janitor")
install.packages("dplyr")
install.packages("skimr")
```

## Load Test Dataset Into Memory

```{r}
library(readxl)
library(janitor)
library(dplyr)
library(skimr)


test_file <- "../../test-data.xlsx"

loans_raw <- read_xlsx(test_file, sheet = 1L) |>
    clean_names()
cashflows_raw <- read_xlsx(test_file, sheet = 2L) |>
    clean_names()
```

## Inspect Data

```{r}
skim(loans_raw)
```

```{r}
skim(cashflows_raw)
```

## Task 1 - Restructure Loan Data

The client has provided us with the loan data in a format that is not compatible
with our internal loan data schema. In particular, the data provided by the
client is a single table that contains the characteristics for both the loans
and the borrowers.
Our internal schema specifies that the loan and borrower details should belong
in separate tables with an additional relationship table linking them both based
on their respective primary ids.

The desired schema for the loan information is described below, where each
key-value pair represents the field name-type belonging to each table

```{r}
target_loan_schema <- list(
    loans = list(
        "contract_number" = "character", # PRIMARY KEY
        "termination_date" = "date",
        "termination_reason" = "character",
        "interest_rate" = "double",
        "loan_amount" = "double",
        "sum_of_last_payment" = "double",
        "debt_principal" = "double",
        "debt_interest" = "double",
        "total_outstanding_and_debt_amount" = "double",
        "proceeding_start_date" = "date",
        "delay_days_principal" = "integer"
    ),
    borrowers = list(
        "customer_id" = "character", # PRIMARY KEY
        "year_of_birth" = "integer",
        "gender" = "character",
        "number_of_customer_active_contracts" = "integer",
        "customer_active_contract_numbers" = "character"
    ),
    id_loan_borrower = list(
        "contract_number" = "character", # PRIMARY KEY
        "customer_id" = "character"
    )
)
```

## Task 1 - Solution

```{r}
# INSERT CODE HERE
```


## Task 2 - Restructure Cashflow Date

Similar to the loan data, the client has provided the cash flow information
in a format that is incompatible with out internal cash flow schema.
Specifically, the client has provided the data in a "wide" format where each
cash flow date is a column. We require the data to be in a "long" format
according to the following schema

```{r}
target_cashflow_schema <- list(
    cashflows = list(
        "contract_number" = "character", # PRIMARY KEY - Component 1
        "cashflow_date" = "date", # PRIMARY KEY - Component 2
        "cashflow_amount" = "double"
    )
)
```

## Task 2 - Solution
```{r}
# INSERT CODE HERE
```


## Task 3 - Calulated Fields

Although the client data set contains all of the data that we require, it is
lacking certain fields that are needed by subsequent data processing pipelines.
In particular, we need to generate the following fields based on the existing
fields we have available:

1. Cumulative Cash Flow Amounts - For each loan we need to have a field that
   contains the sum of all cash flow amounts received to date.

2. Days Past Due (Buckets) - For each loan we require a field that classifies
   the loan based on whether the field `delay_days_principal` falls into one of
   the following buckets:

            [0, 180), [180, 360), [360, 540), [540, 720), 720+

## Task 3 - Solution
```{r}
# INSERT CODE HERE
```


## Task 4 - Automated Data Ingestion Pipeline (Design Question)

*Note: This task is only for candidates applying for the AI Data Engineering
position. If you are applying for a different role, you may skip this task.*

A client sends updated loan and cash flow files on a weekly basis via email.
In practice, the files arrive in varying formats (Excel, CSV, and occasionally
with extra columns, renamed headers, or inconsistent date formats). Currently,
a team member manually inspects and transforms each file using the steps you
completed in Tasks 1-3 before loading it into our internal systems.

We would like to automate this workflow. Design an end-to-end pipeline that
automatically ingests, validates, and transforms incoming client files into the
schemas defined in Tasks 1 and 2. Your design should address the following:

1. **Architecture Overview** - Provide a high-level description (or diagram) of
   the pipeline, from file arrival to data being available in the internal
   schema. Include the key components and how they interact.

2. **Schema Drift Handling** - How would you handle files where columns have
   been added, removed, or renamed compared to what the pipeline expects?

3. **AI/LLM Integration** - Identify where in the pipeline an LLM or AI agent
   could add value (e.g., intelligent column mapping, anomaly detection,
   validation error summarisation, or generating human-readable data quality
   reports). Be specific about what the AI component would receive as input and
   produce as output.

*Note: This is a design question. You are not expected to build a working
pipeline. Focus on demonstrating your thinking around system design, practical
AI integration, and operational reliability.*

## Task 4 - Solution
```{r}
# INSERT YOUR DESIGN / PSEUDOCODE / NOTES HERE
```

## Task 5 - Risk Segmentation and Cash Flow Forecasting Framework

*Note: This task is for candidates applying to the modelling / predictive
analytics projects. It is not expected to require substantial coding. A
concise written answer, table, or simple diagram is sufficient.*

Using the loan and cash flow data from Tasks 1-3, design a simple framework to
support loan valuation, borrower restructuring, or recovery forecasting.

Your response should address the following:

### Objective
Choose one of the following goals:

- forecast future cash flows for a loan,
- segment loans by expected recovery behaviour,
- support restructuring decisions for borrowers in distress.

State what the model or framework is trying to predict or optimise.

### Feature Design
Based on the fields available in the dataset and the derived fields from Tasks
1-3, identify:

- 3-5 input variables you would use,
- why each variable may be predictive,
- any additional derived variables you would create.

### Methodology Proposal
Propose an initial modelling or analytical approach. This does not need to be
mathematically detailed, but should explain:

- what type of approach you would start with (for example: segmentation rules,
  regression, time-series style cash flow projection, survival/default style
  model, or optimisation-based decisioning),
- why that approach is appropriate given the data,
- the key assumptions and limitations.

### Validation and Decision Use
Explain how you would judge whether the framework is useful in practice.
Include:

- one or two evaluation metrics,
- key risks or failure modes,
- how the output would be used by an analyst or decision-maker.

You may answer in bullet points, a short memo, a flow diagram, or a one-page
slide style format. Keep your answer concise.

## Task 5 - Solution
```{r}
# INSERT YOUR RESPONSE HERE
```
